### 2012-07-18 paper2/paper2.tex
another round of changes - We now have a complete draft

% gjilguid2e.tex
% V2.0 released 1998 December 18
% V2.1 released 2003 October 7 -- Gregor Hutton, updated the web address for the style files.

%\documentclass[extra,mreferee]{gji}
\documentclass[extra]{gji}

\bibliographystyle{gji}
\usepackage{timet}
\usepackage[dvips]{graphicx}

\title{Relocating a cluster of earthquakes using a single station}
\author[D.J. Robinson, M. Sambridge, R. Snieder and J Hauser]
  {David J. Robinson$^{1,2}$, Malcolm Sambridge$^{1}$, Roel Snieder$^{3}$, J\"urg Hauser$^{4}$ \\
  $^1$ Research School of Earth Sciences, Australian National
    University, Canberra ACT \emph{0200}, Australia, E-mail: david.robinson@anu.edu.au \\
$^2$  Risk Analysis Methods, Geoscience Australia, GPO Box \emph{383} Canberra ACT \emph{2601} Australia, E-mail: david.robinson@ga.gov.au\\
$^3$ Center for Wave Phenomena and Department of Geophysics, Colorado School of Mines, Golden CO \emph{80401-1887}, USA \\
$^4$ Earthquakes and the Environment Program, NORSAR,Kjeller, Norway}

\date{Received 2012 Month Day; in original form 2012 Month Day}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\volume{200}
\pubyear{2012}

%\def\LaTeX{L\kern-.36em\raise.3ex\hbox{{\small A}}\kern-.15em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
%\def\LATeX{L\kern-.36em\raise.3ex\hbox{{\Large A}}\kern-.15em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
% Authors with AMS fonts and mssymb.tex can comment out the following
% line to get the correct symbol for Geophysical Journal International.
\let\leqslant=\leq

\newtheorem{theorem}{Theorem}[section]

\begin{document}

\label{firstpage}

\maketitle

\begin{summary}
 Coda wave interferometry (CWI) is an emerging tool for constraining 
earthquake source properties from the intereference pattern of coda waves between nearby
events. A new earthquake location algorithm is derived which relies on CWI based probabilistic estimates 
of earthquake separation. The algorithm can be used with CWI estimates alone or in tandem with travel time
data. Synthetic examples in 2D and 3D and real earthquakes on the 
Calaveras Fault, California are used to demonstrate the usefulness of the CWI technique in locating 
poorly recorded earthquakes. It is demonstrated that CWI: (a) outperforms traditional earthquake location 
techniques when the number of stations is small; (b) is self-consistent across a broad range of station
situations; and (c) can be used with a single station to locate earthquakes.

\end{summary}

%\begin{keywords}
% Probability distributions; Earthquake source observations; 
%Seismicity and tectonics; Wave scattering and diffraction
%\end{keywords}

\section{Introduction}

Earthquake location is important for many applications. Locations 
are required for:  magnitude
determination \citep{dr_Richter35a, dr_Gutenberg45a};
computing moment tensors \citep{dr_Sipkin02a};
seismological studies of the Earth's interior 
\citep{dr_Spencer80a, dr_Kennett95a, dr_Curtis02a, dr_Kennett04a};
understanding strong motion and seismic attenuation
\citep{dr_Toro97a, dr_Campbell03a}
 and modeling earthquake hazard or risk
\citep{dr_Frankel00a, dr_Stirling02a, dr_Robinson06b}.

The accuracy required in earthquake location depends on the
application. For example, imaging the structure of a fracture system
from microseismicity requires greater detail than determining
whether a $M_w=7.5$ earthquake occurs offshore for tsunami warning.
 This paper focuses on reducing location uncertainty for 
a cluster of events when they are recorded by a small
number of stations.

Absolute location describes the location of an earthquake with
respect to a global reference such as latitude, longitude (or
easting/northing) and depth. Uncertainties associated with absolute
locations are influenced by source to station distances, the number
of stations and their geometry, quality of signal and accuracy of
the velocity model used in computing travel times. Uncertainties in
absolute location are typically of the order of several km because
they are susceptible to uncertainty in the velocity structure along
the entire path between the source and receiver. For example,
\citet{dr_Shearer99a} states that location uncertainty in the ISC
and PDE catalogues are generally around 25\,km horizontally and at
least 25\,km in depth.\footnote{Here the depth uncertainties of
25\,km assume the use of depth dependent phases such as $pP$.
Without such phases the uncertainty is higher.}
\citet{dr_Bondar04a} demonstrate that at the local scale, absolute
locations are accurate to within 5\,km with a 95\% confidence level
when local networks meet a number of station related criteria. Such
errors are too large for many applications, particularly those
focussed on imaging rupture surfaces from aftershock sequences.

Relative earthquake location involves locating a group of
earthquakes with respect to one another and was first introduced by
\citet{dr_Douglas67a} who developed the technique commonly known as
joint hypocenter determination.\footnote{\citet{dr_Douglas67a}
originally used the term joint epicentre determination. However, he
was solving for hypocentre.} In principle, relative locations can be
computed by differencing absolute locations. However,
\citet{dr_Pavlis92a} shows that inadequate knowledge of velocity
structure leads to systematic biases when relative positions are
computed in this way. To reduce errors from unknown velocity
structure, relative location techniques compute locations directly
from travel time differences between two waveforms 
\citep{dr_Ito85a, dr_Got94a, dr_Nadeau97a, dr_Waldhauser99a}. By doing
so, they remove errors associated with velocity variations outside
the local region, because such variations influence all waveforms in
the same manner \citep{dr_Shearer99a}.

Reported location uncertainties from relative techniques are  around
15 to 75\,m in local settings with good station coverage
\citep{dr_Ito85a, dr_Got94a,
dr_Waldhauser99a,dr_Waldhauser08a}. Here, `good coverage' implies
multiple stations distributed across a broad range of azimuthal
directions. Relative location techniques have been applied to a
diverse range of problems including: location of active fault planes
\citep{dr_Deichmann92a, dr_Got94a, dr_Waldhauser99a,
dr_Waldhauser02a, dr_Shearer05a}; studying rupture mechanics
\citep{dr_Rubin99a, dr_Rubin02a}; interpreting magma
movement in volcanoes \citep{dr_Fremont87a}; and
monitoring pumping induced seismicity \citep{dr_Lees98a,
dr_Ake05a}.

In traditional approaches to absolute and relative location
only early onset body waves, typically $P$ and/or $S$ waves, are used.
The data utilised may be the direct arrival times; the travel time difference computed
between direct arrivals of two waveforms; or the time differences
inferred from time-lagged cross correlation of relatively small windows
around the body wave arrivals.
 In all three cases, the majority of the
waveform is discarded. Furthermore, obtaining high accuracy with these techniques
requires multiple stations with good azimuthal coverage.
In this paper we demonstrate that it is possible to
significantly reduce location uncertainty when few stations are
available by using more of the waveform.

Coda refers to  later arriving waves in the seismogram that arise from
scattering  \citep{dr_Aki69a,dr_Snieder99a,dr_Snieder06a}.
Typically, the coda waves are ignored in most seismological
applications due to the complexity involved in modeling their
generation. In this
paper we develop an approach for locating earthquakes using coda waves.

\citet{dr_Snieder05a} demonstrate that the coda of two earthquakes
can be used to estimate the separation between them. Their
technique, known as coda wave interferometry (CWI), is based on the
interference pattern between the coda waves. Unlike travel time
techniques, CWI
 does not require multiple stations or good azimuthal coverage.
In fact, it is possible to obtain estimates of separation using a single
station \citep{dr_Robinson07b}. This makes CWI particularly interesting
for regions where station density is low such as intraplate settings. In this 
paper we demonstrate how CWI separation estimates can be used to constrain 
location with data from a single station. 
Our technique can be used on coda waves alone or in combination with
travel times.

\section{Theory}
\label{sec:theory}

\citet{dr_Snieder05a} introduce a CWI based estimator of source
separation $\delta_{CWI}$ between two earthquakes
\begin{equation}
\label{eq:Theory-dcwi_g_sigmat}
\delta_{CWI}^2 = g(\alpha,\beta)\sigma_\tau^2,
\end{equation}
where $\sigma_\tau$ is the standard deviation of the travel time perturbation
between the coda waves of two earthquakes,
and $\alpha$ and $\beta$ are the near source $P$ and $S$ wave velocities,
respectively. The function $g$ depends on the type of excitation
(explosion, point force, double couple) and on the direction of
source displacement relative to the point force or double
couple. For example, for two double couple sources displaced in the
fault plane,
\begin{equation}
\label{eq:LitReview-g-3d-dc}
g(\alpha,\beta) = 7
\frac{\left(\frac{2}{\alpha^6}+\frac{3}{\beta^6}\right)}{\left(\frac{6}{\alpha^8}+\frac{7}{\beta^8}\right)},
\end{equation}
whereas, for two point sources in a 2D acoustic medium
\begin{equation}
\label{eq:LitReview-g-2d-ps} g(\alpha,\beta) = 2 \alpha^2
\end{equation}
\citep{dr_Snieder05a}.

The $\sigma_\tau$ in eq. (\ref{eq:Theory-dcwi_g_sigmat}) is related
to $R_{max}$, the maximum of the cross correlation between the coda
of the two waveforms and hence can be computed directly from the
recorded data. In this paper we use the autocorrelation approach of
\citet{dr_Robinson11a} when relating these parameters and we apply a
restricted time lag search when evaluating $R_{max}$. These
extensions to the original technique of \citet{dr_Snieder05a}
increase the range of applicability of CWI by 50\%.

\citet{dr_Robinson11a} show that coda waves provide only
probabilistic constraints on source separation and introduce a
Bayesian approach for describing the probability of true separation
given the CWI data. Their approach is summarised by
\begin{equation}
\label{eq:-bayesian-noisyform}
P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN}) \propto P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)
\times P(\widetilde{\delta}_t)
\end{equation}
where  $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ is the posterior function
indicating the probability of true separation $\widetilde{\delta}_t$ given the noisy CWI separation
estimates $\widetilde{\delta}_{CWIN}$; $P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$
is the likelihood function representing the forward model between the
desired separation $\widetilde{\delta}_t$ and the data $\widetilde{\delta}_{CWIN}$; and
$P(\widetilde{\delta}_t)$ is the prior function accounting for all a-priori information.
The tilde above the separation parameters in eq. (\ref{eq:-bayesian-noisyform})
indicates the use of a wavelength normalised separation parameter
\begin{equation}
\label{eq-normalisation-eqn}
\widetilde{\delta} = \frac{\delta}{\lambda_{d}},
\end{equation}
which measures separation ($\delta$ = $\delta_{CWIN}$ or $\delta_t$)
with respect to dominant wavelength $\lambda_{d}$. In this paper we
consider a uniform prior which gives emphasis to the recorded data.
The algorithm for computing the noisy likelihood
$P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$ is derived by
\citet{dr_Robinson11a} and summarised in Appendix
\ref{sec-Appendix-noisylikelihood}. With these two pieces in place
we can compute the posterior
$P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ (or PDF) for the
separation between any pair of events directly from their coda
waves.

We seek a probability density function (PDF) which links
 individual pairwise posteriors $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$
to describe the location of multiple events with maximum identifying
 the most probable combination of locations.
More importantly however, the PDF will
quantify location uncertainty and provide information
on the degree to which individual events are constrained
 by the data.

For convenience, we begin with three earthquakes having locations
$\mathbf{e}_1$, $\mathbf{e}_2$ and $\mathbf{e}_3$. Using a Bayesian
formulation we write
\begin{equation}
\label{eq-Bayes-three-locations}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3| \mathbf{d}) \propto P( \mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)
\times P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3),
\end{equation}
where $P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \mathbf{d})$,
$P(\mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ and
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ are the posterior,
likelihood and prior functions, respectively. In eq.
(\ref{eq-Bayes-three-locations}) $\mathbf{d}$ represents
observations that constrain the locations. They can be any
combination of travel times, geodetic information  or CWI
separations. For example, if coda waves are used we have
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 |
\widetilde{\mitbf{\delta}}_{CWIN})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN}|\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3)$ where $\widetilde{\mitbf{\delta}}_{CWIN}$ are the
wavelength normalised separation estimates. Alternatively, if we use
CWI and travel time data we may write $P(\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN},
\mathbf{\Delta}_{TT})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN},\mathbf{\Delta}_{TT}|\mathbf{e}_1,
\mathbf{e}_2, \mathbf{e}_3)$ where $\mathbf{\Delta}_{TT}$ represent
travel time differences. In this derivation and in Sec.
\ref{sec:benchmarking} we focus on the constraints imposed by coda
waves, whereas in Sec. \ref{sec:CalaverasLoc-CWIandTT} we
demonstrate how CWI and travel time data can be combined.

For three earthquakes we have likelihoods;
$P(\widetilde{\delta}_{CWIN,12}|\mathbf{e}_1, \mathbf{e}_2)$,
$P(\widetilde{\delta}_{CWIN,13}|\mathbf{e}_1, \mathbf{e}_3)$
and
$P(\widetilde{\delta}_{CWIN,23}|\mathbf{e}_2, \mathbf{e}_3)$.
In writing
these likelihoods we have replaced the conditional term on separation
$\widetilde{\delta}_t$ with the locations (e.g. $\mathbf{e}_1$ and $\mathbf{e}_2$).
This can be done because
knowledge of location translates to separation.\footnote{Note that the reverse
can not be said. That is, knowledge of separation between a single event pair
does not translate to location but rather places a non-unique constraint on
location.}
Furthermore, since the pairwise functions are independent the joint
likelihood becomes
\begin{equation}
\begin{array}{l}
\label{eq:-joint-likelihood4three}
P(\widetilde{\mathbf{\delta}}_{CWIN} | \mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3) =
P(\widetilde{\delta}_{CWIN,12} | \mathbf{e}_1, \mathbf{e}_2) \\
\hspace{2em} \times P(\widetilde{\delta}_{CWIN,13} | \mathbf{e}_1, \mathbf{e}_3)
\times  P(\widetilde{\delta}_{CWIN,23} | \mathbf{e}_2, \mathbf{e}_3).
\end{array}
\end{equation}
Similarly, the earthquake locations are independent and the
joint prior becomes
\begin{equation}
\label{eq:-joint-prior4three}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3) = P((\mathbf{e}_1) \times P(\mathbf{e}_2) \times P(\mathbf{e}_3).
\end{equation}
Combining eqs. (\ref{eq:-joint-likelihood4three}) and (\ref{eq:-joint-prior4three})
gives the joint posterior function
\begin{equation}
\label{eq-3eqposterior}
\begin{array}{l}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN}) = c \displaystyle \prod_{i=1}^3 P(\mathbf{e}_i) \\
\hspace{4em}  \times \displaystyle \prod_{i=1}^2 \prod_{j=i+1}^3 P(\widetilde{\delta}_{CWIN,ij}|\mathbf{e}_i,\mathbf{e}_j)
\end{array}
\end{equation}
for three events.

A detailed understanding of the location of a single event (e.g. $\mathbf{e}_2$) is obtained by computing the
marginal
\begin{equation}
\label{eq:-E2-marginal}
P(\mathbf{e}_2) = \int \int P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \widetilde{\mathbf{\delta}}_{CWIN}) d\mathbf{e}_1 d\mathbf{e}_3
\end{equation}
where the intergral is taken over all plausible locations for $\mathbf{e}_1$ and $\mathbf{e}_3$.
Alternatively, we can compute the marginal for a single event coordinate
by integrating the posterior over all events and remaining coordinates
for the chosen earthquake. Evaluation of $c$ in eq. (\ref{eq-3eqposterior}) involves finding the value which
normalises the posterior function to unit area over all plausible locations.
In many applications the constant of
proportionality $c$ is not important. For example, it is not required
when seeking the combination of locations which maximise
 the posterior function, nor in Bayesian sampling algorithms such as Markov-chain Monte-Carlo
techniques which require evaluation of a function proportional to the PDF.

Following the same argument we get the posterior function
\begin{equation}
\begin{array}{l}
\label{posterior-n-events}
P(\mathbf{e}_1,...,\mathbf{e}_n | \widetilde{\mathbf{\delta}}_{CWIN}) = c \displaystyle \prod_{i=1}^n P(\mathbf{e}_i) \\
\hspace{4em} \displaystyle  \times \prod_{i=1}^{n-1} \prod_{j=i+1}^n P(\widetilde{\delta}_{CWIN,ij}|\mathbf{e}_i,\mathbf{e}_j)
\end{array}
\end{equation}
for $n$ events. When evaluating eq. (\ref{posterior-n-events}) over a range of locations
it is necessary to compute and multiply many numbers close to zero.
This is because the PDFs tend to zero as
the locations get less likely (i.e. near the boundaries of the plausible region).
Such calculations are prone to truncation errors so we
work with the negative logarithm
\begin{equation}
\label{eq:-negative-log-part1}
L(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n) = -ln\left[ P(\mathbf{e}_1,...,\mathbf{e}_n | \widetilde{\mathbf{\delta}}_{CWIN} )  \right]
\end{equation}
or
\begin{equation}
\begin{array}{l}
\label{eq:-negative-log}
L(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n) =
-ln\left[ c \right] - \displaystyle \sum_{i=1}^n ln\left[P(\mathbf{e}_i)\right] \\
\hspace{3em} - \displaystyle \sum_{i=1}^{n-1} \sum_{j=i+1}^n ln\left[P(\widetilde{\delta}_{CWIN,ij}|\mathbf{e}_i,\mathbf{e}_j)\right].
\end{array}
\end{equation}
The logarithm improves numerical stability by replacing products with summations. The negative assists with
optimisation, many algorithms for which are designed to minimise an objective function.

The event locations $\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n$
are defined by coordinates $\hat{x}$, $\hat{y}$ and $\hat{z}$
where the hat $\hat{.}$ indicates use of
a local coordinate system. We choose a local coordinate system
which removes ambiguity associated with transformations of the coordinate system.
In defining this coordinate system we fix the first event at the origin,
\begin{equation}
\label{eq:coord-const1}
\mathbf{e}_1 = (0,0,0),
\end{equation}
the second event on the positive $\hat{x}$-axis
\begin{equation}
\label{eq:coord-const2}
\mathbf{e}_2 = (\hat{x}_2, 0, 0), \hat{x}_2>0
\end{equation}
the third on the $\hat{x}-\hat{y}$ plane
\begin{equation}
\label{eq:coord-const3}
\mathbf{e}_3 = (\hat{x}_3,\hat{y}_3,0), \hat{y}_3>0
\end{equation}
and the fourth to
\begin{equation}
\label{eq:coord-const4}
\mathbf{e}_4 = (\hat{x}_4,\hat{y}_4,\hat{z}_4), \hat{z}_4>0.
\end{equation}
This coordinate system reduces translational (constraint \ref{eq:coord-const1})
and rotational (constraints \ref{eq:coord-const2} to \ref{eq:coord-const4}) non-uniqueness
without loss of generality. It is necessary to work with a local coordinate
system when using coda waves alone because the CWI technique constrains only event separation
between earthquakes. The inclusion of travel times in Section \ref{sec:CalaverasLoc-CWIandTT}
allows us to move to a global reference system.

In summary, the posterior $P(\mathbf{e}_1,...,\mathbf{e}_n | \widetilde{\mathbf{\delta}}_{CWIN})$ and its negative
logarithm $L$ describe the joint probability of multiple event locations given the observed coda waves.
The most likely set of locations is given by the minimum of $L$.   In this paper we use the Polak-Ribiere
technique \citep{dr_Press87a} to optimise $L$. The Polak-Ribiere
technique is a particular choice of conjugate gradient method. It uses the
derivatives of $L$ to guide the optimisation procedure. We derive the derivatives in Appendix \ref{sec-Appendix-derivatives_ofL}.

\section{Synthetic experiments}
\label{sec:benchmarking}

We use synthetic examples in 2D and 3D with 50 earthquakes to test
the performance of the optimisation routine.

\subsection{Examples 1 and 2 - 2D synthetic experiments}

We design a 2D synthetic experiment (Example 1) by
randomly selecting $\hat{x}$- and $\hat{y}$-coordinates such that $-50 \leq \hat{x} \leq 50$ and $-50 \leq \hat{y} \leq 50$.
These are indicated with triangles in Fig. \ref{fig-2D50eq-relocation_eg1}.
 We assume a local velocity of $v=3300\,$ms$^{-1}$ between all
event pairs and a dominant frequency of 2.5$\,$Hz to represent waveform data filtered between 1 and 5$\,$Hz.
The CWI data are defined by the positive bounded Gaussian with statistics
$\bar{\mu}_N$ and $\bar{\sigma}_N$. A hypothetical CWI mean is created by setting
\begin{equation}
\label{eq:hypothetical-CWI-mean-optichapt}
\bar{\mu}_N = \mu_1 \left( \widetilde{\delta}_t \right)
\end{equation}
using eq. (\ref{eq:mu1}). This assumption ensures that the sample
mean of hypothetical separation estimates is consistent with known
CWI biases \citep{dr_Robinson11a}. In example 1 we use
$\bar{\sigma}_N=0.02$ between all event pairs. Application of our
optimisation procedure on the hypothetical CWI data yields the
circles in Fig. \ref{fig-2D50eq-relocation_eg1}. The optimisation
does not lead to the correct solution due to the addition of noise
($\bar{\sigma}_N=0.02$) on the hypothetical CWI data. The average
coordinate error is 2.0\,m which is small compared to the noise of
$\bar{\sigma}_N=0.02$ which for $v_s=3300$\,ms$^{-1}$ and
$f_{dom}=2.5$\,Hz corresponds to roughly 25\,m.
\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_1.eps}
\caption{Example 1 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N=0.02$. Actual and
optimisation event locations are shown in triangles and circles,
respectively.} \label{fig-2D50eq-relocation_eg1}
\end{figure}

\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_3.eps}
\caption{Example 3 - Synthetic relocation of 50 earthquakes in 2D using all constraints with noise
$\bar{\sigma}_N= 2 \sigma_{1to5Hz}(\delta_t)$.
 Actual and optimisation event locations
are shown in triangles and circles, respectively.}
\label{fig-2D50eq-relocation_eg3}
\end{figure}
\citet{dr_Robinson11a} demonstrates that the noise on CWI estimates
is often larger than 0.02 and that it increases with event
separation. Consequently, example 1 is overly simplistic because we
fix $\bar{\sigma}_N=0.02$ for all pairs. In example 2 we increase
the uncertainty and introduce a distance dependance into the
hypothetical $\bar{\sigma}_N$ by defining
$\bar{\sigma}_N=\sigma_{1to5Hz}(\delta_t)$, where
$\sigma_{1to5Hz}(\delta_t)$ is the half-width of the errorbars for a
synthetic acoustic experiment with filtering between 1 and 5\,Hz
\citep[see Fig4(b) of ][]{dr_Robinson11a}. Repeating the
optimisation leads to the circles in Fig.
\ref{fig-2D50eq-relocation_eg3} which have an average coordinate
error of 2.8\,m.

Conjugate gradient based optimisation techniques are susceptible to
the presence of local minima. This is because they use the slope of
the target function to explore the solution space. We explore the
impact of local minima for our CWI location problem by beginning the
optimisation from 25 randomly chosen starting positions. We observe
no differences in the solution for either example.

We can draw three observations from the error structure in Figs.
\ref{fig-2D50eq-relocation_eg1} and \ref{fig-2D50eq-relocation_eg3}.
Firstly, the location errors depicted by grey bars increase between
examples 1 and 2 with the introduction of larger noise. Secondly,
the errors are larger for events at greater distances from the
center. This is because events near the center of the cluster are
constrained by links from all angles, whereas those on the outside
are moderated by links from a limited number of directions. This
observation is analogous to problems associated with poor azimuthal
coverage in common triangulation problems such as individual
earthquake location from limited travel time data, or GPS
positioning with few satellites. Our third observation is that the
location errors appear to be circular, despite our attempt to
correct for rotational non-uniqueness with the local coordinate
system.

The local coordinate system works by constraining the location of
the first three earthquakes. Earthquake 1 is fixed at the origin,
earthquake 2 on the positive $\hat{x}$-axis and earthquake 3 has
$\hat{y}>0$. As the number of events increase the strength of these
constraints on later events weakens allowing small rotations
of events with respect to each other. That is, even though the
rotational freedom of the cluster is in principal removed by the
constraints imposed on the events (see equations (\ref{eq:coord-const1}) to
(\ref{eq:coord-const4})) we observe that in practice the presence of
noise allows the rotational non-uniqueness to reappear. This is
because the `easiest' way data noise can propagate is into the
direction which is least constrained by the data. The same phenomena
is observed in linear inversion where noise creates large spurious
model changes in directions of the eigenvectors with the smallest
singular values \citep{dr_Aster05a}. Unfortunately, there is
no obvious way to overcome this issue when using coda waves alone
since the constraints are based on separation. Fortunately however,
combining coda waves with measurements of travel times alleviates
this problem and facilitate the removal of a local coordinate system
altogether.

Despite our observations about the error structure however, we gain
confidence in the optimisation procedure due to its stability for
different starting locations and because of the small average
coordinate errors of 2.0\,m and 2.8\,m for examples 1 and 2,
respectively.

\subsection{Example 3 - The impact of reduced linkage}

A set of earthquakes and their corresponding coda wave estimates of
separation form a network. In this network the earthquakes are nodes
and the constraints are branches or links which join the nodes
together. Synthetic examples 1 and 2 use 100\% direct linkage
between event pairs. That is, there is a constraint between each
earthquake and all other events. In reality, we might expect that
the separation between some pairs will not be constrained by CWI
data due to poor signal to noise ratio in the coda for common
stations. Obviously, the less stations that record an event the more
likely it is that links between it and other events will be broken.
In such cases the probabilistic distance constraint between a pair
of events may only exist indirectly through multiple pairs. In this
section we consider the impact of reduced linkage between event
pairs. In example 3, we repeat example 2 using 90\%, 80\%, ..., 10\%
of the links. As with the above examples, we undertake the
optimisation with 25 randomly chosen starting locations.

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\noindent\includegraphics[width = 20pc]{diags/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\caption{Statistical measures of error in the optimisation solutions for the 2D synthetic cases when all (grey)
and best (black) results are considered. The statistics $\Delta_{max}$, $\Delta_\mu$ and
$\Delta_\sigma$ are the maximum, mean and standard deviation of the coordinate error, respectively.
The bottom subplot shows the average minimum number of branches required to link the 2450 pairs.}
 \label{fig:optimisationresults-2Dsynth}
\end{figure}

Figure \ref{fig:optimisationresults-2Dsynth}(a) and (b) illustrates 
the maximum $\Delta_{max}$ (top) and mean $\Delta_\mu$ (middle) of the
coordinate difference as a function of increasing linkage. We show
the statistics for the `best' optimisation solution (black) and for
the solution space when all 25 optimisations are considered (grey).
In this case the best solution is determined by the set of event
locations which lead to the smallest value of $L$. The
error in the best solution is consistent when 30\%
or more of the branches are used. The errors increase when only
10\% or 20\% of the constraints are included. Interestingly, this
breakdown around 20\% to 30\% coincides with the point where the
average number of branches required to link each pair reaches 2 (see
Fig.\ref{fig:optimisationresults-2Dsynth} (bottom)). Since the average
number of branches can be computed in advance it can be used as an
indication of the inversion stability prior to
optimisation. An earlier breakdown is observed when all 25 solutions
are considered collectively. For example, the maximum coordinate
error $\Delta_{max}$ exceeds that for the best
solution for linkage $\leq$60\% confirming that the optimisation is susceptible to
local minima and that a range of starting points should
be considered.

Some optimisations 
fail to converge after 1200 iterations when the linkage is 60\% or lower. All optimisations
fail when then linkage is 20\% or lower. Despite their failure to converge however, the locations
at final iteration are close to the actual solution. The derivatives used in the conjugate gradient method
depend on events connected by CWI measurements. Consequently,
earthquakes that are only connected via other events do not `communicate' with each other
directly. To some extent, this should be adressed during the iterative process where location information
can spread to events which have no direct links. However, the
lack of direct connection through the gradient could prevent convergence in extreme cases, or more likely
slow the procedure down. This could explain why some
examples do not converge after 1200 iterations.
\citet{dr_VanDecar94a} show that gradient damping acts
slowly through iterative least-squares, because
 every cell in one iteration communicates only with its neighbours, and they demonstrate that this can be
fixed with preconditioning in some cases. Their findings suggest that it may be possible to improve
the convergence (stability and/or speed) of the CWI optimisation by preconditioning.

\subsection{Example 4 - 3D synthetic examples with reduced linkage}

In Example 4 we expand the optimisation routine to 3D by randomly picking
 a set of actual event locations for 50 earthquakes with
$-50$\,m$\leq \hat{x},\hat{y},\hat{z} \leq 50$\,m. As in the 2D case we assume a local velocity
of $v=3300\,$ms$^{-1}$ between all event pairs and a dominant frequency of 2.5$\,$Hz to represent
 waveform data filtered between 1 and 5$\,$Hz.
The hypothetical CWI mean is created using equation (\ref{eq:hypothetical-CWI-mean-optichapt})
which ensures consistency between the sample mean of hypothetical separation estimates and CWI
biases. We use a standard deviation for the noisy CWI estimates
of $\bar{\sigma}_N = \sigma_{1to5Hz}$ and perform the optimisation using 10\%,
20\%, ..., 100\% of the direct links.
In each case we repeat the
optimisation 25 times using randomly chosen starting locations. The results are
summarised in Figure \ref{fig:optimisationresults-3Dsynth}.

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\noindent\includegraphics[width = 20pc]{diags/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\caption{Statistical measures of error in the optimisation solutions for the 3D synthetic cases when all (grey)
and best (black) results are considered. The statistics $\Delta_{max}$, $\Delta_\mu$ and
$\Delta_\sigma$ are the maximum, mean and standard deviation of the coordinate error, respectively.
The absence of the blue and red lines below 60\% and 30\% indicates a breakdown in the solutions
when all or best optimisation result(s) are considered, respectively.}
\label{fig:optimisationresults-3Dsynth}
\end{figure}

When 70\% of the direct constraints are considered all optimisation
results (grey) are consistent with the best solution (black). The
best solution constrains the event locations down to 30\% of the
direct links.

There is one notable difference between the 3D and 2D results. In 2D
the final iteration was typically close to the actual solution when
the optimisation failed to converge.
Contrastingly, in 3D the optimisation appears to converge to the
correct solution or fail completely, leading to a set locations
which in no way resemble the actual solution. This is depicted in
Figure \ref{fig:optimisationresults-3Dsynth} by the absence of the
grey and black lines below 60\% and 30\% of the constraints,
respectively. The reason for this difference may be due to the
increased number of degrees of freedom in 3D requiring a greater
number of iterations to converge. Nevertheless, the accurate
convergence of the best solution for cases with 30\% linkage or
higher is encouraging for the potential of coda wave optimisation to
constrain earthquake location.

\subsubsection{Summary of synthetic experiments}

In summary, the synthetic examples demonstrate the ability of coda
wave data to constrain relative event location using optimisation.
The optimisation error is influenced by the noise on CWI estimates
with greater $\bar{\sigma}_N$ leading to larger errors in the
solutions. When 70\% or more of the direct branches are used the
optimiser is stable with no observable difference in the solution
for 25 randomly chosen starting locations. As the direct linkage
reduces to 50\% the optimisation becomes less stable and the best
solution from 25 random starting locations is required to find the
optimal solution. As the number of links decrease below
30\% the best solution fails to converge.

\section{Relocating Earthquakes on the Calaveras Fault}
\label{sec:CalaverasLoc-CWIonly}

In this section we relocate 68 earthquakes from the Calaveras Fault,
California. The 68 earthquakes are selected from the 308 earthquake
Calaveras example released with the open source Double Difference
algorithm or hypoDD \citep{dr_Waldhauser00a,dr_Waldhauser01a}. These
events are chosen for three reasons. Firstly, they are recorded by a
large number of stations (Fig.
\ref{fig:-eqopti-California-Calaveras}) and therefore lend
themselves to accurate travel time location. Secondly, they are
distributed over distances from near zero to hundreds
meters making them ideal for application of CWI. Finally, Calaveras
earthquakes have been well researched with several studies having
relocated events in the region \citep{dr_Waldhauser01a,
dr_Schaff02a, dr_Waldhauser08a}. The relocations in this paper are
sorted into four examples summarised in table \ref{tab:examples}.

\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasMap/gmt_california/CaliforniaCalaverasMap1.eps}
\caption{Elevation in California showing location of the 308 event
Calaveras cluster (black star) and 805 seismic stations (red
triangles).} \label{fig:-eqopti-California-Calaveras}
\end{figure}

\begin{table}
\caption{Experiment locations performed for 68 earthquakes on teh
Calaveras Fault} \label{tab:examples}
\begin{tabular}{ll}
\hline
Example 5 & comparison of CWI, catalogue and hypoDD \\
 & locations. \\
Example 6 & exploration of station dependance for CWI and \\
 & hypoDD. \\
Example 7 & combined use of CWI and travel time data \\
& with all and a reduced number of stations. \\
Example 8 & combined use of CWI and travel time data \\
 & when travel times constrain only 50\% of the events. \\
 \hline
\end{tabular}
\end{table}

\subsection{Example 5 - comparison of CWI, catalogue and hypoDD locations}

Figure \ref{fig-69Calaverasevents_eg1} illustrates three sets of
locations for the Calaveras earthquakes. The first column shows the
original catalogue locations for all 308 earthquakes. That is, each event
is located individually using all available travel time arrivals and
a regional velocity model. The 68 earthquakes of interest in this
study are differentiated in black. Catalogue locations suggest that
the 68 earthquakes of interest are distributed spatially throughout
all events.

%%=======================================================================================
% This is the original figure using HypoDD with LSQR
%\begin{figure*}
%\includegraphics{diags/CalaverasLoc1.eps}
%\caption{Comparison of earthquake hypocenters using three different methods: catalogue location (column 1), hypoDD with all
%308 events (column 2) and CWI example 1 (column 3).
%Note that in the case of the CWI locations we consider only the 68 earthquakes in red, the
%blue events are shown for the purpose of orientation only.}
%\label{fig-69Calaverasevents_eg1}
%\end{figure*}

% Now we have the SVD version
\begin{figure*}
\includegraphics{diags/CalaverasLoc1_hypoDD_SVD.eps}
\caption{Comparison of earthquake hypocenters using three different
methods: catalogue location (column 1), hypoDD with all 308 events
(column 2) and CWI example 1 (column 3). Note that in the case of
the CWI locations we consider only the 68 earthquakes in red, the
blue events are shown for the purpose of orientation only.}
\label{fig-69Calaverasevents_eg1}
\end{figure*}
%%=======================================================================================

To apply CWI we download available waveforms 
from the Northern California Earthquake Data Center
(NCEDC). Unsuitable waveforms are removed
using the conditions summarised in Table 5 of
\citet{dr_Robinson11a}. Remaining waveforms are filtered between
1 and 5\,Hz and aligned  to $P$ arrivals at 0\,s. CWI estimates are
obtained from 5\,s wide non-overlapping time windows between $2.5
\leq t \leq 20$\,s and used to create probabilistic constraints on
event separation. We introduce the local coordinate system of
Section \ref{sec:theory} and find the optimum relative locations using 
Polak-Ribiere optimisation.

CWI locations for the 68 events are illustrated in column two of
Figure \ref{fig-69Calaverasevents_eg1}. Catalogue locations (grey)
are shown for the remaining 240 earthquakes and are included to ease
comparison. The third column of Figure
\ref{fig-69Calaverasevents_eg1} illustrates the locations given by
hypoDD with Singular Value Decomposition (SVD), absolute arrival
times and cross correlation computed time differences.

Global locations can not be found by CWI alone. 
For the sake of comparison, we arbitrary choose a
`master' event and translate our relative locations to align with the hypoDD
location for the same event. We return to this issue of relative
versus actual location in Example 7 by introducing a combined travel
time and coda wave inversion.

The spatial distribution of the CWI locations is clearly tighter
than the catalogue locations of column 1. That is, CWI provides an
independent indication of clustering for the 68 events and to first
order, similar locations to those from hypoDD (column 3). Figure 
\ref{fig-69Calaverasevents_eg1}
illustrates a small second order difference between
the CWI and hypoDD based locations. In particular, the lineation is
less clear in the CWI locations (column 2) than that given by
hypoDD (column 3). Our results suggest that the coda are less
supportive of the presence of streaks although a complete
understanding of these differences is left for future work. Our
attention now is devoted towards understanding how both techniques
perform with fewer stations (Example 6) and exploring how CWI and 
travel times can be combined (Examples 7 and 8).

\subsection{Example 6 - Dependance on the number of stations}

Accurate location of the Calaveras events is possible using arrival
phases because of the excellent recording situation in California
with many stations and strong azimuthal coverage (see figure
\ref{fig:-eqopti-California-Calaveras}). In contrast, a small number
of stations and poor azimuthal coverage are common limitations when
trying to locate intraplate clusters. For example, there are
only 3 network seismic stations in the South West Seismic Zone of
Western Australia, a region similar in size to that hosting 805
stations in Figure \ref{fig:-eqopti-California-Calaveras}.

We explore the impact of poorer recording situations in Example 6 by
re-locating the 68 Calaveras events  using hypoDD and coda waves
with a reduced number of stations. We begin with 10 stations and
repeat the process removing one at a time until a single station
remains. The 10 stations considered are shown in Figure
\ref{fig:-eqopti-Calaveras-substations} and the order of removal
explained in Table \ref{tab:Calaveras-stationremoval}.

\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasMap/matlab/Calaveras_substationmap}
\caption{Location of the 10 stations (triangles) used to relocate
the Calaveras events in Example 2. Stations are removed one at a
time according to the order in Table
\ref{tab:Calaveras-stationremoval} and the events relocated. The
events are indicated with black circles.}
\label{fig:-eqopti-Calaveras-substations}
\end{figure}

\begin{table}
\caption{Stations considered when exploring the imnpact of reduced
station coverage.} \label{tab:Calaveras-stationremoval}
\begin{tabular}{|r|l|}
\hline
Number of & Station Names\\
Stations  & \\
\hline
10 & CCO, JCB, JST, CMH, HSP, JAL, CSC, JST, CAD, JHL, JRR\\
9  & CCO, JCB, JST, CMH, HSP, JAL, CSC, JST, CAD, JHL\\
8  & CCO, JCB, JST, CMH, HSP, JAL, CSC, JST, CAD\\
7  & CCO, JCB, JST, CMH, HSP, JAL, CSC \\
6  & CCO, JCB, JST, CMH, HSP, JAL \\
5  & CCO, JCB, JST, CMH, HSP \\
4  & CCO, JCB, JST, CMH \\
3  & CCO, JCB, JST \\
2  & CCO, JCB \\
1  & CCO \\
\hline
\end{tabular}
\end{table}

CWI locations are illustrated in Figure
\ref{fig-CWIreducesstats} for the inversions with 7, 5, 4, 3, 2 and
1 station. We observe a high level of consistency between these 6
inversions and the locations shown in Figure
\ref{fig-69Calaverasevents_eg1} (column 2) when all stations are
considered. That is, the coda wave approach is self-consistent
regardless of the number of stations available, reinforcing our
hypothesis that coda waves can constrain location in poor recording
situations.

\begin{figure*}
\includegraphics[angle=90,height = 50pc]{diags/CalaverasLoc2.eps}
\caption{....} \label{fig-CWIreducesstats}
\end{figure*}

Figure \ref{fig-HYPODDreducesstats} illustrates the hypoDD inversion
results for 7, 5 and 4 stations. The travel time problem is
ill-posed for fewer than 4 stations so it is not possible to apply
hypoDD with SVD for 3 or fewer stations.  The hypoDD locations are
not self-consistent as the number of stations is reduced. We observe
a general increase in scatter and a higher number of stray events
outside the cluster when less stations are used with hypoDD. Even
with 7 stations the linear streak of Figure
\ref{fig-69Calaverasevents_eg1} (column 3) is less evident.

%%=========================================================================
% this is the original version using hypoDD with LSQ
%\begin{figure*}
%\includegraphics[angle=90,height = 50pc]{diags/CalaverasLoc3.eps}
%\caption{....}
%\label{fig-HYPODDreducesstats}
%\end{figure*}

% This is the version using hypoDD with SVD
\begin{figure*}
\includegraphics[height = 25pc]{diags/CalaverasLoc3_hypoDD_SVD.eps}
\caption{HypoDD (SVD) with reduced stations.}
\label{fig-HYPODDreducesstats}
\end{figure*}
%%=========================================================================

As the number of stations are reduced both the CWI and hypoDD
techniques are not able to re-locate all events. To use the coda
waves we need at least one pairwise separation constraint to be
formed from the available stations. This means that for every event
there must be at least one station that records it and at least one
other earthquake sufficiently well to apply CWI. Fortunately, we can
make an assessment of this prior to starting the inversion. The top
row of Figure \ref{fig-statremoval_summarystats} demonstrates that
when 5 or more stations are used CWI can constrain the location of
all 68 earthquakes. When less than 5 stations are used the coda
waves constrain a decreasing number of events until at 1 station it
is only possible to locate 55 events.

%%=========================================================================
% This is the original version using hypoDD with LSQ
%\begin{figure}
%\noindent\includegraphics[width = 20pc]{diags/CalaverasLoc4.eps}
%\caption{Statistics on coordinate differences for reduced station inversions. Differences
%are computed between the inversion results (CWI and hypoDD) and the complete hypoDD
%locations for all 308 events. The top subplot illustrates the number of constrainable
%events in the CWI and hypoDD inversions as a function of the stations considered.}
%\label{fig-statremoval_summarystats}
%\end{figure}

% This is the version using hypoDD with SVD
\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasLoc4_hypoDD_SVD.eps} \caption{Statistics on
coordinate differences for reduced station inversions. Differences
are computed between the inversion results (CWI and hypoDD) and the
complete hypoDD locations for all 308 events. The top subplot
illustrates the number of constrainable events in the CWI and hypoDD
inversions as a function of the stations considered.}
\label{fig-statremoval_summarystats}
\end{figure}
%%=========================================================================

The hypoDD program also struggles to locate all events as the number
of stations is reduced. In the case of hypoDD an event can be
identified as unconstrainable in one of two stages. Firstly, the
data are analysed to ensure that there exists travel time
differences for each event and at least one other earthquake. This
is analogous to the situation for the coda wave technique. The
hypoDD program also has a secondary identification phase in which
events that can not be located sufficiently are rejected during the
inversion. This process is presumably related to the iterative
removal of outliers described by \citet{dr_Waldhauser00a}. The top
row of Figure \ref{fig-statremoval_summarystats} shows that the
number of events re-located by hypoDD fluctuates between 63 and 28
earthquakes for 10 to 4 stations and it demonstrates that the number
of events located by hypoDD is less than the number located by CWI
in all cases.

The remaining rows of Figure \ref{fig-statremoval_summarystats}
illustrate a statistical comparison of the CWI and hypoDD reduced
station locations to those using hypoDD with all available data. For
the CWI inversions the mean and maximum coordinate difference is
consistent regardless of the number of stations considered. In
contrast, the statistics for the hypoDD inversions fluctuate between
... and .... for the mean coordinate error and ... and ...  for the
maximum confirming that the hypoDD inversion procedure is less
stable than CWI with fewer stations.

\section{Theory for Combining travel time and CWI constraints}
\label{sec:CalaverasLoc-CWIandTT}
In Examples 5 and 6 we compare the location of the Calaveras
earthquakes using coda wave and arrival time based constraints
independently. Since the arrival time (direct or difference) and
coda wave data
 come from different sections of the waveform they provide independent constraints on the locations.
 In this section we devise a location algorithm which incorporates
 both CWI and travel time data.

We do not propose a new technique for earthquake location using
travel time differences. Rather, we exploit the information created
by hypoDD with SVD to define a probability density (or posterior)
function
\begin{equation}
\label{eq-multi-var-Gauss-tt} 
\begin{array}{l}
P(\mathbf{e}_p|\Delta_{TT}) 
\frac{1}{(2\pi)^{\frac{3}{2}}\sqrt{|\Sigma|}} \\
\hspace{5em} \times \exp
\left({-\frac{1}{2}\left([\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T
\Sigma^{-1} [\mathbf{e}_p-\mu_{\mathbf{e}_p}]\right)} \right),
\end{array}
\end{equation}
where
\begin{equation}
\mathbf{e}_p = (x_p,y_p,z_p)^T
\end{equation}
is the location of event $p$,
\begin{equation}
\mu_{\mathbf{e}_p} = (\mu_{x_p}, \mu_{y_p},\mu_{z_p})^T
\end{equation}
is the most likely location as determined using the travel time
data, and
\begin{equation}
\label{eq:Sigma-expression}
\Sigma = \left( \begin{array}{ccc} \sigma_{x_p}^2 & 0 & 0\\
0 &  \sigma_{y_p}^2 & 0 \\
0 & 0 & \sigma_{z_p}^2  \end{array} \right)
\end{equation}
is the covariance matrix. In this paper we define the mean location
$\mu_{\mathbf{e}_p}$ and covariance matrix by the hypoDD optimum
solution and its uncertainties. It is important to note that hypoDD
must be used with SVD to obtain useful estimates of $\sigma_{x_p}$,
$\sigma_{y_p}$ and $\sigma_{z_p}$ (\textbf{ADD REF}).

We pose the location problem using the negative log likelihood
\begin{equation}
\label{eq-Lstar-tt-cwi}
\begin{array}{l}
 L(\mathbf{e}_1, \mathbf{e}_2, ...,
\mathbf{e}_1, \mathbf{e}_n) = - \sum_{i=1}^n
ln\left[P(\mathbf{e}_i|\Delta_{TT})\right] \\
\hspace{6em}  - \sum_{i=1}^{n-1}
\sum_{j=i+1}^n
ln\left[P(\delta_{CWIN}|\mathbf{e}_i,\mathbf{e}_j)\right],
\end{array}
\end{equation}
where $(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n)$ is the joint
location,
\begin{equation}
\label{eq-ttcomponent} \sum_{i=1}^n
ln\left[P(\mathbf{e}_i|\Delta_{TT})\right]
\end{equation}
incorporates the travel time constraints and
\begin{equation}
\label{eq-codacomponent} \sum_{i=1}^{n-1} \sum_{j=i+1}^n
ln\left[P(\delta_{CWIN}|\mathbf{e}_i,\mathbf{e}_j)\right]
\end{equation}
 the coda waves.

We must differentiate $L$ to use the Polak-Ribiere conjugate
gradient technique of \citet{dr_Press87a}. The derivative of
$L(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n)$ with respect to
$x_p$ is given by
\begin{equation}
\begin{array}{l}
\frac{\partial L}{\partial x_p} =  - \frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial x_p}
- \sum_{i=p+1}^{N} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_p,\mathbf{e}_i)\right]}{\partial
x_p} \\
\hspace{5em} - \sum_{j=1}^{p-1} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_j,\mathbf{e}_p)\right]}{\partial
x_p}
\end{array}
\end{equation}
where 
\begin{equation}
\sum_{i=p+1}^{N} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_p,\mathbf{e}_i)\right]}{\partial
x_p} 
\end{equation}
and 
\begin{equation}
\sum_{j=1}^{p-1} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_j,\mathbf{e}_p)\right]}{\partial
x_p} 
\end{equation}
are defined in Appendix \ref{sec-Appendix-derivatives_ofL} and
\begin{equation}
\begin{array}{l}
\frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial x_p}
= -\frac{1}{2}[1,0,0]^T \Sigma^{-1}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}] \\
\hspace{9em} -\frac{1}{2}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T \Sigma^{-1} [1,0,0].
\end{array}
\end{equation}
Similarly, for the derivatives with respect to $y_p$ and $z_p$ we
have
\begin{equation}
\begin{array}{l}
\frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial y_p}
= -\frac{1}{2} [0,1,0]^T \Sigma^{-1}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}] \\
\hspace{9em} -\frac{1}{2}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T \Sigma^{-1} [0,1,0] 
\end{array}
\end{equation}
and
\begin{equation}
\begin{array}{l}
\frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial z_p}
= -\frac{1}{2} [0,0,1]^T \Sigma^{-1}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}] \\
\hspace{9em} -\frac{1}{2}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T \Sigma^{-1} [0,0,1].
\end{array}
\end{equation}

Aggregating the travel time and coda wave data offers two
advantages. Firstly, it combines independent constraints on the
event locations offering further confidence in the resulting
solution. Secondly, the travel time constraints in the form of
equation (\ref{eq-ttcomponent}) resolve the inherent non-uniqueness
associated with translation, rotation and reflection around a global
coordinate system. This means that it is no longer necessary to use
a local coordinate system and we can solve directly for location
with respect to a global reference. Collectively, these advantages
improve the behavior of the Polak-Ribiere optimisation leading to
faster and more stable convergence. Consequently, we no longer have
to consider multiple randomly chosen starting locations.

\subsection{Example 7 - Combining travel time and CWI constraints}
 Figure
\ref{fig-68Calaverasevents_ttandcoda1} illustrates the earthquake
locations obtained when we combine the travel time and coda wave
data using all data (left) and 5 stations (right). The linear
features observed in the original hypoDD inversions (see figure
\ref{fig-69Calaverasevents_eg1}) is evident in both cases. However,
the coda waves introduce a scatter around these streaks. That is,
the locations in figure \ref{fig-68Calaverasevents_ttandcoda1}
result from a trade-off between hypoDD's desire to place the events
on linear features and the coda waves voracity to push them away
from streaks. When all stations are used the hypoDD constraints are
strong and little off-streak scatter is introduced. As we reduce
hypoDD's leverage by decreasing the number of stations to 5, we
observe an increase in off-streak scatter resulting from the
enhanced influence of the coda.

%%=========================================================================
% This is the original version using hypoDD with LSQ
% This is only makes sense for LSQR so we remove it from this manuscript
%\begin{figure*}
%\noindent\includegraphics{diags/CalaverasLoc5.eps}
%\caption{Locations of the 68 Calaveras earthquakes combining all available coda wave and travel time constraints with three different
%levels of uncertainty on the travel time PDFs. Column 1 sets $\sigma_x = \sigma_y = 19.5$\,m and $\sigma_z = 15$\,m
%\citep[after]{dr_Waldhauser08a}, Column 2 uses $\sigma_x = \sigma_y = 3 \times19.5$\,m and $\sigma_z = 3\times15$\,m and
%column 3 considers $\sigma_x = \sigma_y = 152$\,m and $\sigma_z = 232$\,m \citep[after]{dr_Shearer97a}. }
%\label{fig-68Calaverasevents_ttandcoda1}
%\end{figure*}

% This is the version using hypoDD with SVD
\begin{figure}
\includegraphics{diags/CalaverasLoc5_hypoDD_SVD.eps}
\caption{Combined HypoDD (SVD) and CWI: all stations and 5
stations.} \label{fig-68Calaverasevents_ttandcoda1}
\end{figure}

%%=========================================================================

\subsection{Example 4 - Combining CWI and travel times when the travel times constrain a limited number of events}

In intraplate regions such as Australia it is common to deploy
temporary seismometers to monitor aftershocks for significant events
\citep{dr_Bowman90a, dr_Leonard02a}). Traditionally, these
deployments are used to locate aftershocks to an accuracy of the
temporary array, whereas events occurring outside the deployment
period can only be located to the accuracy of the network. Using our
combined inversion it is possible to re-locate all events by
employing the detailed travel time data when the temporary network
is in-situ and using coda waves from network stations when the
deployment is absent.

An estimate of the cummulative number of aftershocks $N(t)$ after 
$t$ days is given by the modified Omori formula
\begin{equation}
N(t) = K \frac{c^{1-p} + (t+c)^{1-p}}{p-1}
\end{equation}
\citep{dr_Utsu95a}.
The empirically derived constants, $K$, $C$ and $p$ vary between
tectonic settings. For example, using recorded aftershocks with
$M\ge3.2$ of the  Hokkaido-Nansei-Oki, Japan $M_s=7.8$ earthquake of
12 July 1993, \citet{dr_Utsu95a} obtained maximum likelihood
estimates for $K$, $p$ and $c$ of 906.5, 1.256 and 1.433,
respectively.  With these empirically derived values an array
deployed within 4 days and left for 150 days will record roughly one
half of the aftershocks occurring within the first 1000 days. That
is,
\begin{equation}
\frac{N(150+4)-N(4)}{N(1000)} = \frac{2257-934}{2626} \approx 0.5.
\end{equation}
This idea is illustrated in Figure \ref{fig:Omorifigure} which shows
the best fitting Omori Formula separated into segments before
(dashed), during (solid) and after (dashed) the pseudo temporary
deployment.

\begin{figure}
\noindent\includegraphics[width = 20pc]{diags/OmoriFigure.eps}
\caption{Cumulative number of aftershocks for the
Hokkaido-Nansei-Oki, Japan $M_s=7.8$ earthquake of 12 July 1993
according to the best fitting modified Omori Formula
\citep{dr_Utsu95a}. The leftmost dashed, middle solid and rightmost
dashed signify aftershocks occurring before, during and after the
deployment of a temporary array installed 4 days after the main
shock and left for 150 days. A temporary deployment of this kind
will record roughly 50\% of the aftershocks in the 1000 days
following the mainshock. } \label{fig:Omorifigure}
\end{figure}

With this idea of a temporary deployment in mind we have another
attempt at relocating the Calaveras earthquakes. In Example 8
we consider the travel time constraints on half (34) of the
earthquakes and incorporate coda wave data from a single stations
for all 68 earthquakes. The combined inversion is shown in column 1
of Figure \ref{fig-68Calaverasevents_ttsubsetandcoda1}. The
inversion result is similar to the combined inversion when all
travel time data is incorporated (see Fig.
\ref{fig-68Calaverasevents_ttandcoda1}). The slight increase in
scatter observed here can be explained by the events with no travel
time constraints and the tendency of the coda to push events away
from streaks.

Remarkably, the combined coda wave and travel time inversion locates
all 68 earthquakes to an accuracy similar to the inversions with all
data. In contrast when travel time data is used alone it is only
possible to locate the 34 events recorded by the pseudo temporary
deployment. This
ability of  coda waves to constrain the location of events recorded
by a single station creates new opportunities for understanding
earthquakes in regions with limited station coverage.

%%=========================================================================
% This is the original version using hypoDD with LSQ
%\begin{figure}
%\noindent\includegraphics{diags/CalaverasLoc6.eps}
%\caption{Left - Combined travel time and coda wave inversions using travel time constraints on
%22 (or 32\%) of the events and coda waves from station CCO only. All travel time based uncertainties
%are assigned $\sigma_x = \sigma_y = 3 \times19.5$\,m and $\sigma_z = 3\times15$\,m. Right - travel time
%relocation only (i.e. no CWI) using ........}
%\label{fig-68Calaverasevents_ttsubsetandcoda1}
%\end{figure}

% This is the version using hypoDD with SVD
\begin{figure}
\includegraphics{diags/CalaverasLoc6_hypoDD_SVD.eps}
\caption{Mimicking the deployment of a temporary network by ignoring
data from all but station CCO for 50\% (or 34) of the events using
by combining CWI and TT (left) and using TT only (right). Only by
combining the data is it possible to locate all 68 events.
Furthermore, combining the data leads to a solution more consistent
with figure ..... } \label{fig-68Calaverasevents_ttsubsetandcoda1}
\end{figure}
%%=========================================================================

\section{Discussion and Conclusions}

Coda wave interferometry is an emerging technique for constraining 
earthquake location. The technique relies on the interference between coda waves
of closely located events and is hence most useful
for studying earthquake clusters and/or aftershock sequences. 
Coda wave constraints are independent of travel times
and can be used in isolation or combination with early onset body waves. The strength of coda
is that it is possible to constrain earthquake location from a single station, 
an outcome demonstrated most clearly by Figures .......

Coda wave interferometry offers a new technique for understanding earthquakes in intraplate areas 
with sparse networks and poor azimuthal coverage.
In particular, the ability to combine coda wave constraints with travel times makes it possible
to link well constrained events from 
a temporary deployment with those recorded outside
the deployment period. All that is required to achieve this is at 
least one network station which has recorded
sufficient events from both periods therefore opening new avenues
for imaging intraplate fault structures and 
improving our understanding of intraplate seismicity and earthquake hazard.

Another potential application of CWI is in the area of hydraulic fracturing 
such as hot rock geothermal projects and/or
petroleum reservoir engineering. Monitoring pumping induced micro earthquakes is a key step in understanding
the migration of fluids in such reservoirs. There is a trade-off in the ability of surface deployed networks
to locate events which are small and/or deep. As we begin to push reservoir projects to deeper depths
downhole seismic monitoring is likely to play increasingly important roles. 
CWI creates new possibilities to monitor 
pumping induced micro earthquakes from fewer boreholes and
hence dramatically reduce the costs of reservoir monitoring at large depths. It may also be possible
to utilise coda for understanding hazard in tunneled mining operations where the 
location of deep tunnels prohibits
azimuthal coverage of induced events.

\begin{acknowledgments}
Geoscience Australia, the Research School of Earth Sciences at The
Australian National University, and the Center for Wave Phenomena at
the Colorado School of Mines, are acknowledged for supporting this
research. The paper is published with permission of the CEO
Geoscience Australia. Work was conducted as part of an Australian
Research Council Discovery Project (DP0665111). We thank the
Northern California Earthquake Data Center (NCEDC) for providing the
Calaveras data and the Northern California Seismic Network (NCSN),
U.S. Geological Survey, Menlo Park and Berkeley Seismological
Laboratory, University of California, Berkeley for contributing it
to the NCEDC. We also acknowledge Felix Waldhauser and William
Ellsworth, the authors of the openly available Double Difference
location algorithm, hypoDD. This paper has benefited significantly
from reviews by ......... at Geoscience Australia.
\end{acknowledgments}

\bibliography{../refs/drrefs}

\appendix

\section{The noisy likelihood}
\label{sec-Appendix-noisylikelihood}

The noisy likelihood $P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$
used in eq. (\ref{eq:-bayesian-noisyform}) is given by
\begin{equation}
\begin{array}{l}
\label{eq-likelihood-int}
P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)  =
A(\widetilde{\delta}_t) C(\bar{\mu}_N, \bar{\sigma}_N)  \\
\hspace{3em} \times \int_0^\infty
B(\widetilde{\delta}_t,\widetilde{\delta}_{CWI})
D(\widetilde{\delta}_{CWI},\bar{\sigma}_N,\bar{\mu}_N )
d\widetilde{\delta}_{CWI}
\end{array}
\end{equation}
where $\widetilde{\delta}_{CWI}$ is an estimate of CWI separation in the absence
of noise,
\begin{equation}
\label{eq:Adefn}
A(\widetilde{\delta}_t) = \frac{1}{(1-\Phi_{\mu_1,\sigma_1}(0))\sigma_1\sqrt{2\pi} },
\end{equation}
\begin{equation}
B(\widetilde{\delta}_t,\widetilde{\delta}_{CWI})=e^{  \frac{-(\widetilde{\delta}_{CWI}-\mu_1)^2}{2\sigma_1^2} },
\end{equation}
\begin{equation}
\label{eq:Cdefn}
C(\bar{\mu}_N, \bar{\sigma}_N) =  \frac{1}{(1-\Phi_{\bar{\mu}_N,\bar{\sigma}_N}(0))\sigma_N\sqrt{2\pi}},
\end{equation}
\begin{equation}
D(\widetilde{\delta}_{CWI},\bar{\sigma}_N,\bar{\mu}_N )=e^{  \frac{-(\widetilde{\delta}_{CWI}-\bar{\mu}_N)^2}{2 \bar{\sigma}_N ^2} }
\end{equation}
and $\Phi_{\mu,\sigma}(x)$ is the cumulative Gaussian distribution function
\begin{equation}
\label{eq-cummulative-Gaussian}
\Phi_{\mu,\sigma}(x) = \frac{1}{\sigma \sqrt{2 \pi}}
\int_{-\infty}^x e^{  \frac{-(s-\mu)^2}{2\sigma^2}  } ds
\end{equation}
\citep{dr_Robinson11a}. The parameters $\mu_1$ and $\sigma_1$ used
in eq. (\ref{eq:Adefn}) are defined by the expresions
\begin{equation}
\label{eq:mu1}
\mu_1(\widetilde{\delta}_t) = a_1\frac{a_2 \widetilde{\delta}_t^{a_4}+a_3
\widetilde{\delta}_t^{a_5}}{a_2 \widetilde{\delta}_t^{a_4}+a_3 \widetilde{\delta_t}^{a_5}+1}
\end{equation}
and
\begin{equation}
\label{eq:sigma1}
\sigma_1(\widetilde{\delta}_t) = c + a_1\frac{a_2 \widetilde{\delta}_t^{a_4}+
a_3 \widetilde{\delta}_t^{a_5}}{a_2 \widetilde{\delta}_t^{a_4}+a_3 \widetilde{\delta}_t^{a_5}+1}
\end{equation}
with coefficients $a_1$ to $a_5$ and $c$ defined in Table \ref{tab-const4-mu1-sigma1}.
The parameters $\bar{\mu}_N$ and $\bar{\sigma}_N$ used in
eq. (\ref{eq:Cdefn}) are obtained by finding the values which minimise the
difference in a least squares sense between the noisy CWI estimates $\widetilde{\delta}_{CWIN}$
computed from the waveforms and the
positively bounded Gaussian density function
\begin{equation}
\label{eq-likelihood-noisydata-pdf-orig}
\begin{array}{l}
P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t,\widetilde{\delta}_{CWI}) \\
\hspace{5em} = \frac{1}{\left(1-\Phi_{\bar{\mu}_N,\bar{\sigma}_N}(0)\right)\bar{\sigma}_N\sqrt{2\pi}}
e^{  \frac{-(\widetilde{\delta}_{CWIN}-\bar{\mu}_N)^2}{2\bar{\sigma}_N^2}  }
\end{array}
\end{equation}
with $\widetilde{\delta}_{CWIN} \geq 0$.

\begin{table}
\caption{Coefficients for eqs. (\ref{eq:mu1}) and (\ref{eq:sigma1}).}
\label{tab-const4-mu1-sigma1}
\begin{tabular}{|c|c|}
\hline
$\mu_1(\widetilde{\delta}_t)$ & $\sigma_1(\widetilde{\delta}_t)$ \\
\hline
$a1 = 0.4661$ & $a1 = 0.1441$\\
$a2 = 48.9697$ & $a2 = 101.0376$\\
$a3 = 2.4693$ & $a3 = 120.3864$\\
$a4 = 4.2467$ & $a4 = 2.8430$\\
$a5 = 1.1619$ & $a5 = 6.0823$ \\
     & $c = 0.017$ \\
\hline
\end{tabular}
\end{table}

\section{Derivatives}
\label{sec-Appendix-derivatives_ofL}

The derivatives of $L(\mathbf{e}_1,\mathbf{e}_2,...,\mathbf{e}_N)$
\begin{equation}
\frac{\partial L}{\partial \hat{x}_1},
\frac{\partial L}{\partial \hat{y}_1},
 \frac{\partial L}{\partial \hat{z}_1},
\frac{\partial L}{\partial \hat{x}_2},
\frac{\partial L}{\partial \hat{y}_2},
\frac{\partial L}{\partial \hat{z}_2},
...,
\frac{\partial L}{\partial \hat{x}_N},
\frac{\partial L}{\partial \hat{y}_N},
\frac{\partial L}{\partial \hat{z}_N}
\end{equation}
are required by the Polak-Ribiere algorithm. These are used to guide the optimisation procedure
towards the values of $(\mathbf{e}_1,\mathbf{e}_2,...,\mathbf{e}_N)$ which
minimise $L$.

The equations for the derivatives are convoluted so we
build them gradually. We start with an expression for $\delta_t$, the wavelength normalised separation between two events
$\mathbf{e}_p= (\hat{x}_p,\hat{y}_p,\hat{z}_p)$ and $\mathbf{e}_q = (\hat{x}_q,\hat{y}_q,\hat{z}_q)$
\begin{equation}
\label{eq:deltat4chapt6deriv}
\delta_t = \frac{f_{dom}}{v_s}\sqrt{(\hat{x}_p-\hat{x}_q)^2 + (\hat{y}_p-\hat{y}_q)^2 +  (\hat{z}_p-\hat{z}_q)^2},
\end{equation}
where $f_{dom}$ is the dominant frequency of the waveforms and $v_s$ is the velocity between the events. Expression
\ref{eq:deltat4chapt6deriv} has derivatives
\begin{equation}
\label{eq-partial-xyz1}
\begin{array}{l}
\frac{\partial \widetilde{\delta}_t}{\partial \hat{x}_p} = \frac{f_{dom}^2 (\hat{x}_p-\hat{x}_q)}{v_s^2 \widetilde{\delta}_t},
\frac{\partial \widetilde{\delta}_t}{\partial \hat{y}_p} = \frac{f_{dom}^2 (\hat{y}_p-\hat{y}_q)}{v_s^2 \widetilde{\delta}_t}, \\
\frac{\partial \widetilde{\delta}_t}{\partial \hat{z}_p} = \frac{f_{dom}^2 (\hat{z}_p-\hat{z}_q)}{v_s^2 \widetilde{\delta}_t},
\frac{\partial \widetilde{\delta}_t}{\partial \hat{x}_q} = \frac{f_{dom}^2 (\hat{x}_q-\hat{x}_p)}{v_s^2 \widetilde{\delta}_t}, \\
\frac{\partial \widetilde{\delta}_t}{\partial \hat{y}_q} = \frac{f_{dom}^2 (\hat{y}_q-\hat{y}_p)}{v_s^2 \widetilde{\delta}_t},
\frac{\partial \widetilde{\delta}_t}{\partial \hat{z}_q} = \frac{f_{dom}^2 (\hat{z}_q-\hat{z}_p)}{v_s^2 \widetilde{\delta}_t}.
\end{array}
\end{equation}
For brevity we focus the following derivation in terms of $\hat{x}_p$. The remaining terms for $\mathbf{e}_p$
(i.e. $\hat{y}_p$ and $\hat{z}_p$) can be computed
by following the same procedure. The derivatives for $\mathbf{e}_q$ can be attained by exploiting the symmetry
\begin{equation}
\label{eq:ep2eq-deriv-symmetry}
\frac{\partial \widetilde{\delta}_t}{\partial \hat{x}_q} = - \frac{\partial \widetilde{\delta}_t}{\partial \hat{x}_p}.
\end{equation}

The chain rule gives
\begin{equation}
\frac{\partial \mu_1}{\partial \hat{x}_p} = \frac{\partial \mu_1}{\partial \widetilde{\delta}_t} \frac{\partial \widetilde{\delta}_t}{\partial \hat{x}_p}
\end{equation}
where differentiating eq. (\ref{eq:mu1}) gives
\begin{equation}
\label{eq-dmu1-by-dx1}
\frac{\partial \mu_1}{\partial \widetilde{\delta}_t} = a_1 \frac{a_2 a_4 \widetilde{\delta}_t^{a_4-1} +a_3 a_5 \widetilde{\delta}_t^{a_5-1}}
{\left(a_2 \widetilde{\delta}_t^{a_4} +a_3 \widetilde{\delta}_t^{a_5} +1 \right)^2}.
\end{equation}
Similarly, we have
\begin{equation}
\frac{\partial \sigma_1}{\partial \hat{x}_p} = \frac{\partial \sigma_1}{\partial \widetilde{\delta}_t} \frac{\partial \widetilde{\delta}_t}
{\partial \hat{x}_p}
\end{equation}
where $\frac{\partial \sigma_1}{\partial \widetilde{\delta}_t}$ has the identical form as \ref{eq-dmu1-by-dx1} with different constants
$a_1,a_2,...,a_5$ (see table \ref{tab-const4-mu1-sigma1}).

The cumulative Gaussian distribution function \ref{eq-cummulative-Gaussian} is
\begin{equation}
\Phi_{\mu_1,\sigma_1}(0) = \frac{1}{\sigma_1 \sqrt{2 \pi}}
\int_{-\infty}^0 e^{  \frac{-(s-\mu_1)^2}{2\sigma_1^2}  } ds
\end{equation}
which has derivative
\begin{equation}
\frac{\partial \Phi_{\mu_1,\sigma_1}(0)}{\partial \hat{x}_p} =
\frac{ \sigma_1 \int_{-\infty}^0 \frac{\partial g}{\partial \hat{x}_p} e^g ds -
\frac{\partial \sigma_1}{\partial \hat{x}_p} \int_{-\infty}^0 e^g ds}
{\sigma_1^2 \sqrt{2 \pi}}
\end{equation}
where
\begin{equation}
g = \frac{-(s-\mu_1)^2}{2 \sigma_1^2}
\end{equation}
and
\begin{equation}
\frac{\partial g}{\partial \hat{x}_p} = \frac{4 \sigma_1^2 (s-\mu_1) \frac{\partial \mu_1}{\partial \hat{x}_p}
+ 4\sigma_1 \frac{\partial \sigma_1}{\partial \hat{x}_p}(s-\mu_1)^2}
{4 \sigma_1^4}.
\end{equation}

Now, we have all the pieces to compute the derivatives of $A=A(\delta_t)$ and $B = B(\delta_t,\delta_{CWI})$ as follows
\begin{equation}
\frac{\partial A} {\partial \hat{x}_p} =
-\frac{ -\frac{\partial  \Phi_{\mu_1,\sigma_1}(0)}{\partial \hat{x}_p} \sigma_1 +
 \left(1- \Phi_{\mu_1,\sigma_1}(0) \right) \frac{\partial \sigma_1}{\partial \hat{x}_p}}
{\left(1- \Phi_{\mu_1,\sigma_1}(0) \right)^2 \sigma_1^2 \sqrt{2 \pi}}
\end{equation}
and
\begin{equation}
\frac{\partial B} {\partial \hat{x}_p} = e^h \frac{\partial h}{\partial \hat{x}_p}
\end{equation}
where
\begin{equation}
h =  \frac{-(\delta_{CWI}-\mu_1)^2}{2 \sigma_1^2}
\end{equation}
and
\begin{equation}
\frac{\partial h}{\partial \hat{x}_p} = \frac{
4 \sigma_1^2 (\delta_{CWI}-\mu_1) \frac{\partial \mu_1}{\partial \hat{x}_p}
+ 4(\delta_{CWI}-\mu_1)^2 \sigma_1 \frac{\partial \sigma_1}{\partial \hat{x}_p} }
{4 \sigma_1^4}.
\end{equation}
Finally, we can differentiate the likelihood for an individual event pair
\begin{equation}
\begin{array}{l}
\frac{\partial P(\delta_{CWIN}|\widetilde{\delta}_t)} {\partial \hat{x}_p}  =
\frac{\partial A(\widetilde{\delta}_t)}{\partial \hat{x}_p} C(\bar{\mu}_N, \bar{\sigma}_N) \\
\hspace{2em} \times \int_0^\infty B(\widetilde{\delta}_t,\widetilde{\delta}_{CWI})
D(\widetilde{\delta}_{CWI},\bar{\sigma}_N,\bar{\mu}_N )
d\widetilde{\delta}_{CWI} \\
\hspace{2em} + A(\widetilde{\delta}_t) C(\bar{\mu}_N, \bar{\sigma}_N) \\
\hspace{2em} \times \int_0^\infty
\frac{\partial B(\widetilde{\delta}_t,\widetilde{\delta}_{CWI})} {\partial \hat{x}_p}
D(\widetilde{\delta}_{CWI},\bar{\sigma}_N,\bar{\mu}_N )
d\widetilde{\delta}_{CWI}
\end{array}
\end{equation}
and for the logarithm we have
\begin{equation}
\frac{ \partial \ln \left[ P(\delta_{CWIN}|\delta_t) \right] } {\partial \hat{x}_p}
= \frac{1}{P(\delta_{CWIN}|\delta_t)} \frac{\partial P(\delta_{CWIN}|\delta_t)}{\partial \hat{x}_p}.
\end{equation}

Thus, it follows that the derivative of $L$ with respect to $\hat{x}_p$ is given by
%\begin{equation}
%\label{eq-derivative-Lstar-cwionly22}
%\frac{\partial L(E_1, E_2, ..., E_n)}{\partial \hat{x}_p} =
%- \sum_{i=p+1}^{N} \frac{ \partial \ln \left[P(\delta_{CWIN}|E_p,E_i)\right]}{\partial \hat{x}_p}
%+ \sum_{j=1}^{p-1} \frac{ \partial \ln \left[P(\delta_{CWIN}|E_j,E_p)\right]}{\partial \hat{x}_p}
%\end{equation}
\begin{equation}
\label{eq-derivative-Lstar-cwionly}
\begin{array}{ll}
\frac{\partial L(E_1, E_2, ..., E_n)}{\partial \hat{x}_p} = &
- \sum_{i=p+1}^{N} \frac{ \partial \ln \left[P(\delta_{CWIN}|E_p,E_i)\right]}{\partial \hat{x}_p} \\
 & + \sum_{j=1}^{p-1} \frac{ \partial \ln \left[P(\delta_{CWIN}|E_j,E_p)\right]}{\partial \hat{x}_p}
\end{array}
\end{equation}
%
%\iftwocol{\begin{eqnarray} \frac{\partial L(E_1, E_2, ..., E_n)}{\partial \hat{x}_p} &  = &\\
%- \sum_{i=p+1}^{N} \frac{ \partial \ln \left[P(\delta_{CWIN}|E_p,E_i)\right]}{\partial \hat{x}_p} \\
%+ \sum_{j=1}^{p-1} \frac{ \partial \ln \left[P(\delta_{CWIN}|E_j,E_p)\right]}{\partial \hat{x}_p}
%\end{eqnarray}}
%{\frac{\partial L(E_1, E_2, ..., E_n)}{\partial \hat{x}_p} =
%- \sum_{i=p+1}^{N} \frac{ \partial \ln \left[P(\delta_{CWIN}|E_p,E_i)\right]}{\partial \hat{x}_p}
%+ \sum_{j=1}^{p-1} \frac{ \partial \ln \left[P(\delta_{CWIN}|E_j,E_p)\right]}{\partial \hat{x}_p}}
%\end{equation}
for a uniform prior. The change of sign in the middle (i.e. to addition) accounts for the change in
order of the events under the conditional. Its inclusion here assumes the correct use of
$\partial \widetilde{\delta}_t / \partial \hat{x}_p$
or
$\partial \widetilde{\delta}_t / \partial \hat{x}_q$
when evaluating the left and right hand terms of the summation. The derivatives shown in this section
appear complicated but are in practice trivial to compute numerically.
Confidence in their accuracy is enhanced in the following sections by demonstrating that
the optimisation procedure converges to the correct solution for a number of synthetic problems
in 2 and 3 dimensions.

\clearpage

%\bsp % ``This paper has been produced using the Blackwell
     %   Publishing GJI \LaTeXe\ class file.''

\label{lastpage}

\end{document}